---
title: "Original KC with Disaggregated Error Types"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

```{r}
library(lme4)
library(dplyr)
library(ggplot2)
library(knitr)
library(reshape2)
library(reshape)
```


### Transaction to Rollup

+ Load transaction

```{r}
# load a transaction file annotated with the split errors
read_transaction_file <- function(path){
  df <- read.csv(path)
  dataframe <- data.frame(df, na.string = "", stringsAsFactors = FALSE)
  return(dataframe)
}
  
```


+ Add KC model according to model definition

```{r}
#Use this function if using the new KC model

#df: transaction file,
#new_kc: dataframe which defines better model
#old_kc_model_name: column name of old kc 

add_kc_model <- function(df, new_kc, old_kc_model_name){
  new_df <- left_join(df, new_kc, by = old_kc_model_name)
  return(new_df)
}
```



+ Concatenate four binary columns to binary string
```{r}


concatenate <- function(df){
  # rows where correct
  correct_row <- which((df[, "S_current"] == 1) & (df[, "I_current"] == 1))
  
  # might redefine correct ?? now treat every correct as "1100"
  df[correct_row, ]$S_downstream <- 0
  df[correct_row, ]$I_downstream <- 0
  
  # make new column of binary stirng
  df$binary_string <- paste(as.numeric(df$S_current), as.numeric(df$I_current), 
                            as.numeric(df$S_downstream), as.numeric(df$I_downstream),sep = "")

  return(df)
}

```




+ Add error type from binary string according to error definition
```{r}

add_error_types <- function(df, error_definition){
  
  df <- concatenate(df)
  df <- left_join(df, error_definition, by = "binary_string")

  return(df)
}

```



+ Add hint and correct error to error types
```{r}
#After adding the error types defined in error_definition, add in the hint error type and 
# also correct ones
  
add_hint_correct <- function(df, error_definition){
  colname <- c(colnames(error_definition)[2:3])

#  df[, colname] <- as.character(df[, colname])
  hint_row <- which(df[, "Outcome"] == "HINT")
  df[hint_row, colname[1]] <- "hint"
  df[hint_row, colname[2]] <- "hint"
  
  correct_row <- which(df[, "Outcome"] == "CORRECT")
  df[correct_row, colname[1]] <- "correct"
  df[correct_row, colname[2]] <- "correct"
  return(df)
}

```



+ Make step slices
```{r}
# For every student, for each step,
# Make step slices, each step slice distinguished by sumCorrect
make_step_slices <- function(df){
  
 # drop_col <- c("Session.Id", "Time.Zone", "Duration..sec.", "Student.Response.Type",
#                "Student.Response.Subtype", "Level..Assignment.", "Level..ProblemSet.",
#                "Problem.View","Attempt.At.Step", "Is.Last.Attempt","Feedback.Text",
#                "Feedback.Classification","Help.Level", "Condition.Name", "Condition.Type")
  
  df <- df %>%
   # select (-drop_col) %>% 
    group_by(Anon.Student.Id, Step.Name) %>% 
    mutate(correct = ifelse(Outcome == "CORRECT", 1, 0),
                     sumCorrect = cumsum(correct))
  
  df[df$correct == 0, ]$sumCorrect <- df[df$correct == 0, ]$sumCorrect + 1 
  return(df)
  
}
```



+ Rollup
```{r}
transaction_to_rollup <- function(df, model_definition, model, error_definition, error){
  if(model == "orig"){
    m <- colnames(model_definition)[1]
  }else if(model == "new"){
    m <- colnames(model_definition)[2]
  }
 
  if(error == "orig"){
    e <- colnames(error_definition)[2]
  }else if(error == "new"){
    e <- colnames(error_definition)[3]
  }
  

  df <- df %>% group_by(Anon.Student.Id, Step.Name, sumCorrect) %>% 
               mutate(KC_name =  tail(!!as.name(m), n = 1),  # KC_name is the KC student is working toward,
               first_error = (!!as.name(e))[1],         # first error in the step slice
               hint_used = ifelse("HINT" %in% Outcome, "hint", ""), # Whether has used hint in that step slice
               gen = ifelse( (Outcome[1] == "INCORRECT" | Outcome[1] == "HINT"), 1, 0) # general error
               ) %>% 
                
       distinct(df, .keep_all=TRUE) # Keep only first row for each step slice, so #row = #step slice

  return(df)
}
```



+ Add indicator columns indicating which first error according to error definition 

```{r}

add_first_error <- function(roll){
  unique_type <- unique(roll$first_error)
  unique_type <- unique_type[!is.na(unique_type)]
  for (type in unique_type){
    roll[, type] <- ifelse(roll$first_error == type, 1, 0)
  }
  return(roll)
}

```



+ Add opportunity count
```{r}
# Count opp for each student:
# For each student, for each KC being worked toward, count number of step slice(rows)

add_opportunity <- function(roll){
  
  roll <- roll %>% group_by(Anon.Student.Id, KC_name) %>% 
  mutate(opp = seq.int(n())) #  %>% arrange(by_group = KC_name) 

  roll <- roll[roll$KC_name != "", ]
  roll$opp <- as.factor(roll$opp)
  return(roll)
}


```




+ Aggregate errors
```{r}

# Aggregated across KC
aggregate_all <- function(roll, exclude_KC = ""){
  unique_type <- unique(roll$first_error)
  unique_type <- unique_type[!is.na(unique_type)]
  # Filter out some KC
  
  agg <- roll[!(roll$KC_name %in% exclude_KC), ]

  
  col <- c("opp", "gen")
  for (type in unique_type){
   col <- c(col, type)
  }
  agg <- agg[, (colnames(agg) %in% col)]
  

  # Make 3 aggregated tables
  agg1 <- agg %>% group_by(opp) %>% summarise(n = n())
  agg2 <- agg %>% group_by(opp) %>% summarise_all(.funs = mean, na.rm = T) 
  agg3 <- agg %>% group_by(opp) %>% summarise_all(.fun = sum, na.rm = T)
  return(list(agg1, agg2, agg3))
  
}



# By individual KC
aggregate_kc <- function(roll, exclude_KC = ""){
  unique_type <- unique(roll$first_error)
  unique_type <- unique_type[!is.na(unique_type)]
  # Filter out some KC
  
  agg <- roll[!(roll$KC_name %in% exclude_KC), ]
  
  col <- c("opp", "KC_name", "gen")
  for (type in unique_type){
   col <- c(col, type)
  }
  agg <- agg[, (colnames(agg) %in% col)]

  # Make 3 aggregated tables
  agg1 <- agg %>% group_by(KC_name, opp) %>% summarise(n = n())
  agg2 <- agg %>% group_by(KC_name, opp) %>% summarise_all(.funs = mean, na.rm = T) 
  agg3 <- agg %>% group_by(KC_name, opp) %>% summarise_all(.fun = sum, na.rm = T)
  return(list(agg1, agg2, agg3))
  
}

```


### Plotting

+ Overall plot
```{r}

plot_all <- function(res, exclude_error = "None", y_range, main){
  
  agg <- res[[2]]
  agg$opp <- as.numeric(agg$opp)
  agg <- data.frame(agg)
  err_type <- (colnames(agg))[-1]
  
  if (exclude_error != "None"){
    err_type <- err_type[!(err_type %in% exclude_error)]
  }
  #print(err_type)
  agg <- melt(agg, id = "opp", measure = err_type) %>% filter(variable != "correct")
  ggplot(agg, aes(opp, value, colour = variable)) + geom_line() + ylim(y_range) + xlim(0,25) + labs(title = main, x = "Opportunity", y = "Error Rate")
}  
  


plot_kc <- function(res, exclude_error = "None", y_range, main){
  
  agg <- res[[2]]
  agg$opp <- as.numeric(agg$opp)
  agg <- data.frame(agg)
  err_type <- (colnames(agg))[-c(1,2)]
  print("plot kc, err_type:")
  print(err_type)
  #print(colnames(agg))
  #print()
  
  
  if (exclude_error != "None"){
    err_type <- err_type[!(err_type %in% exclude_error)]
  }
  #print(err_type)
  agg <- melt(agg, id = c("opp", "KC_name"), measure = err_type) %>% filter((variable != "correct") & (variable != "gen" ))
  ggplot(agg, aes(opp, value, colour = variable)) + geom_line() + ylim(y_range) + xlim(0,25) + facet_wrap(~ KC_name) + labs(title = main, x = "Opportunity", y = "Error Rate")
}   
  

```


+ Residual plot
```{r}
residual_plot <- function(tables1, tables2, exclude_error = "None", h_line = FALSE, y_range = c(-0.2, 0.2), main = ""){
  t1 <- tables1[[2]]
  t2 <- tables2[[2]]
  # First col is opp,rest are errors
  # Reorder cols
  order <- colnames(t1)
  t2 <- t2[ ,order]
  
  # Truncate
  nrows <- min(nrow(t1), nrow(t2))
  longer <- which.max(c(nrow(t1), nrow(t2)))
  diff <- abs(nrow(t1) - nrow(t2))
  
  
  print(paste("Truncating", as.character(diff), "rows from table", as.character(longer), "...", sep = " "))
  t1 <- t1[c(1:nrows),]
  t2 <- t2[c(1:nrows),]
  
  residual <- t1 - t2
  
  residual$opp <- 1:nrow(residual)
  residual <- data.frame(residual)
  err_type <- (colnames(residual))[-1]
  
  if (exclude_error != "None"){
    err_type <- err_type[!(err_type %in% exclude_error)]
  }
  
  residual <- melt(residual, id = "opp", measure = err_type) %>% filter(variable != "correct")
  p <- ggplot(residual, aes(opp, value, colour = variable)) + geom_line() + ylim(y_range) + xlim(0,25) + labs(title = main, x = "Opportunity", y = "Value")
  
  if(h_line == TRUE){
    p <- p + geom_hline(aes(yintercept = 0), linetype = "dashed") 
  }
  p
  
  
}

```



### Wrapper functions

```{r}


rollup_from_transaction <- function(transaction, error_definition, model_definition, old_model_name, model_to_use = "orig", error_to_use = "orig"){
  path <- transaction
  
  # clean transaction
  trans <- read_transaction_file(path)
  trans <- add_kc_model(trans, model_definition, old_model_name)
  trans <- add_error_types(trans, error_definition)
  trans <- add_hint_correct(trans, error_definition)
  trans <- make_step_slices(trans)
  
  rollup <- transaction_to_rollup(trans, model_definition, model = model_to_use, error_definition, error = error_to_use)
  rollup <- add_first_error(rollup)
  rollup <- add_opportunity(rollup)
  
  return(rollup)
}


# use this to make residual, returns a list of 3 tables
aggregate_from_rollup <- function(rollup, KC_to_remove = "", mode = "aggregated"){
  if(mode == "aggregated"){
    agg_res <- aggregate_all(rollup, exclude_KC = KC_to_remove)
  }else if(mode == "KC"){
    agg_res <- aggregate_kc(rollup, exclude_KC = KC_to_remove)
  }
  return(agg_res)
}


plot_from_aggregate <- function(table, mode = "aggregated", error_filter = "None", y_range = c(0, 0.5), main){
  if(mode == "aggregated"){
    plot <- plot_all(table, exclude_error = error_filter, y_range = y_range, main = main)
  }else if(mode == "KC"){
    plot <- plot_kc(table, exclude_error = error_filter, y_range = y_range, main = main)
  }
  return(plot)
}


```










### Define better KC model
```{r}

orig_kc <- c("M den5", "M num5", "M done", "AS den5", "AS num5", 
             "AS check_convert", "AS done", "", "AD num3","AD check_convert", 
             "AD den3", "AD done", "AD den4", "AD num4", "AD num5", "AD den5", 
             "M check_convert", "AD operation2", "M blankProblem")


better_kc <-  c("M den_ans", "M num_ans", "done", "A den_ans", 
                "A num_ans", "check_convert", "done", "", "AD num_conv", 
                "check_convert", "AD den_conv", "done", "AD den_conv", 
                "AD num_conv", "A num_ans", "A den_ans", "check_convert", 
                "AD operation2", "M blankProblem")

model_defined <- data.frame(matrix(NA, nrow = length(orig_kc), ncol = 2))
names(model_defined) <- c("KC..Field.", "KC_combined")
model_defined$KC..Field. <- orig_kc
model_defined$KC_combined <- better_kc


kable(model_defined[-8, ]) 
#(model_defined
```




### Define error types from binary string
```{r}
#Binary strings that appear in the transaction file:
#"1100", "1000", "0000", "0011", "1001", "0110", "0010", "0111", "NA"   
#where "NA" means that transaction is either tutor performed, or is a hint request

binary_str <- c("1100", "1000", "1001", "0000", "0010", "0110", "0011", "0111")
type_defined <- data.frame(matrix(NA, nrow = length(binary_str), ncol = 3))
orig_type <- c("correct", "incorrect", "misapplied", "out_of_graph", "wild", "where", "when", "where")

simplified_type <- c("correct", "incorrect",  "incorrect",  "if-error", "incorrect", "if-error", "if-error", "if-error")

names(type_defined) <- c("binary_string", "original_error_type", "simplified_error_type")
type_defined$binary_string <- binary_str
type_defined$original_error_type <- orig_type
type_defined$simplified_error_type <- simplified_type


kable(type_defined)
```










### Load human data

```{r}

human <- rollup_from_transaction(transaction = "/Users/daniellaye/Documents/transactions/trans_human.csv",
                        error_definition = type_defined, 
                        model_definition = model_defined, 
                        old_model_name = "KC..Field.", 
                        model_to_use = "orig", error_to_use = "orig")

```

```{r}
agg_all_without_done_human <- aggregate_from_rollup(rollup = human, KC_to_remove = c("M done", "AS done", "AD done"), mode = "aggregated")
agg_all_with_done_human <- aggregate_from_rollup(rollup = human, mode = "aggregated")

agg_kc_without_done_human <- aggregate_from_rollup(rollup = human, KC_to_remove = c("M done", "AS done", "AD done"), mode = "KC")
agg_kc_with_done_human <- aggregate_from_rollup(rollup = human, mode = "KC")
```


### Load AL data


```{r}
al <- rollup_from_transaction(transaction = "/Users/daniellaye/Documents/transactions/trans_iso.csv",
                        error_definition = type_defined, 
                        model_definition = model_defined, 
                        old_model_name = "KC..Field.", 
                        model_to_use = "orig", error_to_use = "orig")
```


```{r}
agg_all_without_done_al <- aggregate_from_rollup(rollup = al, KC_to_remove = c("M done", "AS done", "AD done"), mode = "aggregated")
agg_all_with_done_al <- aggregate_from_rollup(rollup = al, mode = "aggregated")

agg_kc_without_done_al <- aggregate_from_rollup(rollup = al, KC_to_remove = c("M done", "AS done", "AD done"), mode = "KC")
agg_kc_with_done_al <- aggregate_from_rollup(rollup = al, mode = "KC")
```


## Residuals (AL - Human)

### Without done and without total error

```{r}
# Without done and without total error
r1 = residual_plot(agg_all_without_done_al, agg_all_without_done_human, exclude_error = c("gen"), h_line = TRUE, y_range = c(-0.1, 0.1), main = "Residual Plot without Done and Without Total Error (AL - Human)")
r1
```

### With done and without total error

```{r}
# With done and without total error
r2 = residual_plot(agg_all_with_done_al, agg_all_with_done_human, exclude_error = c("gen"), h_line = TRUE, y_range = c(-0.1, 0.1), main = "Residual Plot with Done and Without Total Error (AL - Human)")
r2
```

### Overall total error without done KC

```{r}
# Overall total error without done KC
r3 = residual_plot(agg_all_without_done_al, agg_all_without_done_human, h_line = TRUE, y_range = c(-0.2, 0.2), main = "Residual Plot without Done and with Total Error (AL - Human)")
r3
```

### Overall total error with done KC

```{r}
# Overall total error with done KC
r4 = residual_plot(agg_all_with_done_al, agg_all_with_done_human, h_line = TRUE, y_range = c(-0.2, 0.2), main = "Residual Plot with Done and with Total Error (AL - Human)")
r4
```



## Tables of overall error rates

### AL With done 

```{r}
# AL With done 
kable(agg_all_with_done_al[[2]])

```


### AL Without done

```{r}
# AL Without done
kable(agg_all_without_done_al[[2]])
```


### Human With done 

```{r}
# Human With done 
kable(agg_all_with_done_human[[2]])

```


### Human Without done

```{r}
# Human Without done
kable(agg_all_without_done_human[[2]])
```





## Overall plots for AL

### Overall without done KC and without total error

```{r}
# Overall without done KC and without total error
al1 <- plot_from_aggregate(agg_all_without_done_al, mode = "aggregated", error_filter = "gen", y_range = c(0, 0.2), main = "Overall without Done and without Total Error for AL")
al1

```


### Overall with done KC and without total error

```{r}
# Overall with done KC and without total error
al2 <- plot_from_aggregate(agg_all_with_done_al, mode = "aggregated", error_filter = "gen", y_range = c(0, 0.2), main = "Overall with Done and without Total Error for AL")
al2

```

### Overall total error without done KC

```{r}
# Overall total error without done KC
al3 <- plot_from_aggregate(agg_all_without_done_al, mode = "aggregated", y_range = c(0, 0.6), main = "Overall without Done and with Total Error for AL")
al3

```


### Overall total error with done KC

```{r}
# Overall total error with done KC
al4 <- plot_from_aggregate(agg_all_with_done_al, mode = "aggregated", y_range = c(0, 0.6), main = "Overall with Done and with Total Error for AL")
al4

```


### KC with done 

```{r}
# KC with done for AL
al5 <- plot_from_aggregate(agg_kc_with_done_al, mode = "KC",  error_filter = "gen", y_range = c(0, 0.8), main = "Error Rate by KC with Done and without Total Error for AL")
al5


```

### KC without done 

```{r}
# KC without done for AL
al6 <- plot_from_aggregate(agg_kc_without_done_al, mode = "KC",  error_filter = "gen", y_range = c(0, 0.8), main = "Error Rate by KC without Done and without Total Error for AL")
al6


```





## Human

### Overall without done KC and without total error

```{r}
# Overall without done KC and without total error
hu1 <- plot_from_aggregate(agg_all_without_done_human, mode = "aggregated", error_filter = "gen", y_range = c(0, 0.2), main = "Overall without Done and without Total Error for Human")
hu1

```

### Overall with done KC and without total error

```{r}
# Overall with done KC and without total error
hu2 <- plot_from_aggregate(agg_all_with_done_human, mode = "aggregated", error_filter = "gen", y_range = c(0, 0.2), main = "Overall with Done and without Total Error for Human")
hu2

```


### Overall total error without done KC

```{r}
# Overall total error without done KC
hu3 <- plot_from_aggregate(agg_all_without_done_human, mode = "aggregated", y_range = c(0, 0.6), main = "Overall without Done and with Total Error for Human")
hu3

```


### Overall total error with done KC

```{r}
# Overall total error with done KC
hu4 <- plot_from_aggregate(agg_all_with_done_human, mode = "aggregated", y_range = c(0, 0.6), main = "Overall with Done and with Total Error for Human")
hu4

```


### KC with done 

```{r}
# KC with done for Human
hu5 <- plot_from_aggregate(agg_kc_with_done_human, mode = "KC",  error_filter = "gen", y_range = c(0, 0.8), main = "Error Rate by KC with Done and without Total Error for Human")
hu5

```

### KC without done 

```{r}
# KC without done for Human
hu6 <- plot_from_aggregate(agg_kc_without_done_human, mode = "KC",  error_filter = "gen", y_range = c(0, 0.8), main = "Error Rate by KC without Done and without Total Error for Human")
hu6


```

